### 2017-08-07 examples/Example_001_ER_CTCF.Rmd
Created simiplified examples of the analyses in the ./examples subdirectory.

---
title: "Normalization Example using internal CTCF peaks as a standard"
author: "Andrew Holding"
date: "8/7/2017"
output: pdf_document
---

## Introduction

This is a normalisation using linear regression to CTCF of MCF7 cells chip treameant etc.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load convience functions

These functions facilitate the normalisation of data.

```{r brundle}
source('../package/brundle.R')
```

## Apply settings

```{r}
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
jg.treatedCondition       =  "Fulvestrant"
jg.untreatedCondition     =  "none"

```

## Load control and experimental DiffBind object

To keep file size down these are provided as a Rdata File rather than as raw counts.

```{r}
filename<-"Rdata/example_001_SLX-14438_dba_human_ER_CTCF.rda"
if(!file.exists(filename)){
  dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates=TRUE)
  dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates=TRUE)
  save(dbaExperiment,dbaControl,file=filename)
} else {
  load(filename)
}

```

```{r}
#Load Sample Ids from control sample sheet.
jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
```

```{r}

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

```

```{r}
#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

```

```{r}
#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

```

```{r}

##Plot showing normalization calculation (Optional)
  jg.plotNormalization(jg.controlCountsTreated,
                     jg.controlCountsUntreated)
```

```{r}
##Get Normalization Coefficient
jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)
jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames)

```

```{r}
#Apply coefficent and control factor
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames)
```

```{r}
#Return values to Diffbind and plot normalised result.
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised)

```

```{r}
jg.dba_analysis<-dba.analyze(jg.dba)
dba.plotMA(jg.dba_analysis,bFlip=TRUE)

```

## Save results

```{r}
write.csv(dba.report(jg.dba_analysis),file="results/Example_001.csv")
```