### 2017-08-07 examples/Example_002_ER_CTCF_DeSeq.Rmd

---
title: "Normalization Example using internal CTCF peaks as a standard (no-linear fit)"
author: "Andrew Holding"
date: "8/7/2017"
output: pdf_document
---

## Introduction

This is a normalisation using only DeSeq Size Factors to CTCF of MCF7 cells chip treameant etc.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load convience functions

These functions facilitate the normalisation of data.

```{r brundle, echo=FALSE}
source('../package/brundle.R')
```

## Apply settings

```{r}
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"

```

## Load control and experimental DiffBind object

To keep file size down these are provided as a Rdata File rather than as raw counts.

```{r}
filename<-"Rdata/example_001_SLX-14438_dba_human_ER_CTCF.rda"
if(!file.exists(filename)){
  dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates=TRUE)
  dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates=TRUE)
  save(dbaExperiment,dbaControl,file=filename)
} else {
  load(filename)
}

```

Diffbind was simply used as a convient way to extract peak counts. The actual DESeq Pipeline starts simply needs a matrix so we covert from Diffbind and save as csv to provide a starting point for a DESeq pipeline.

```{r}

filename<-"csv/experimentalPeakset.csv"
if(!file.exists(filename)){
  ## Extract Peak set from DiffBind
  jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
  #Convert Peakset to DeSeq Workflow
  jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)
  #Save to CSV

  write.csv(jg.experimentPeaksetDeSeq,file=filename)
}

#Repeat for control samples.

filename<-"csv/controlPeakset.csv"
if(!file.exists(filename)){
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)
write.csv(jg.controlPeaksetDeSeq,   file=filename) 
}

```

# Pipe line starts here
## Load CSV dataset in DESeq format

```{r}
jg.controlPeaksetDeSeq <- read.csv( file="csv/controlPeakset.csv"
                                      ,check.names=FALSE, row.names = 1)
jg.experimentPeaksetDeSeq<- read.csv(file="csv/experimentalPeakset.csv"
                                     ,check.names=FALSE, row.names = 1)
```

Note we use the samplesheet here, not needed but neater.

```{r}

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)

```

## Export results as CSV

These are not annotated, but other packages exist to do that.
```{r}

write.csv(file="results/Example_002.csv", jg.experimentResultsDeseq)

```

```{r}

#Plot results

  jg.plotDeSeq(jg.controlResultsDeseq,
               p=0.01, 
               title.main="Fold-change in CTCF binding",
               flip=T
               )

jg.plotDeSeq(jg.experimentResultsDeseq,
             p=0.01,
             title.main="Fold-change in ER binding",
             flip=T
             )

#Draw Combined figure.
  jg.plotDeSeqCombined(jg.controlResultsDeseq,
                       jg.experimentResultsDeseq,
                       title.main="ER and CTCF Binding Folding changes on ER treatment",
                       p=0.01,flip=TRUE)

```