### 2017-08-14 examples/Example_003_ER_H2av.Rmd

---
title: "Normalisation Example using Spike-in Control as a Standard"
author: "Andrew Holding"
date: "7/8/2017"
output: pdf_document
---

## Introduction

This is a normalisation using H2av binding within an external spike-in of drosophilia chromatin to provide a control for the analysis of ER binding in MCF7 cells after treatment with Fulvestrant.

This example builds on the previous two examples and therefore they should be read first as many details of key functions have not been repeated here.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preprocessing

Unlike Examples 001 and 002 the use of a xenogenic spike-in requires more preprocessing. The scripts for this process are provided below. Genomes were downloaded from http://support.illumina.com/sequencing/sequencing_software/igenome.html.

The first step of our preprossing pipeline requires the generation of a combined human/drosophila genome. We used the following script to generate this. The human and drosophila genomes downloaded from the link provided above are contained in a subdirectory "genomes" relative to where the script was run. These example scripts are not evalutated in this markdown due to the time it takes to process the data.

```{bash, eval=FALSE}

# Merge the genomes

human=../Homo_sapiens/UCSC/hg19/Sequence/WholeGenomeFasta/genome.fa
droso=../Drosophila_melanogaster/UCSC/dm3/Sequence/WholeGenomeFasta/genome.fa

mkdir genomes/dmhs
cd ./genomes/dmhs
sed "s/^>/>hs_/" $human > tmp
sed "s/^>/>dm_/" $droso > tmp2
cat tmp tmp2 > dmhs.fa
rm tmp tmp2

# Merge Annotations
human=../Homo_sapiens/UCSC/hg19/Annotation/Archives/archive-current/Genes/genes.gtf
droso=../Drosophila_melanogaster/UCSC/dm3/Annotation/Archives/archive-current/Genes/genes.gtf

sed "s/^/hs_/" $human > tmp
sed "s/^/dm_/" $droso > tmp2

cat tmp tmp2 > dmhs.gtf
rm tmp tmp2

```

To allow us to use the merged genome we first indexed it using bowtie2. For our system, we specified 32 threads; the number of threads will depend on the number of processors you have available.

```{bash, eval=FALSE}
### Index the merged genome
cd ./genomes/dmhs

bowtie2-build --threads 32 dmhs.fa dmhs

```

Once indexed, we could then align to the merged genome. This script assumes that the FastQ files are in a directory ./SLX-8047_dhms from where the scripts are run.

```{bash, eval=FALSE}
#Align, Sort and Build indexes

mkdir ./SLX-8047_dmhs
cd ./SLX-8047_dmhs

genome=../genomes/dmhs/dmhs
for fq in ../SLX-8047/*fq.gz
do
  root=`basename $fq`
  bowtie2 -p 32 -x $genome -U $fq > tmp.sam \
  && samtools view -Sbh tmp.sam > tmp.bam \
  && samtools sort tmp.bam ${root}  \
  && samtools index ${root}.bam
done
rm tmp.sam tmp.bam
```

Blacklisting requires a merged blacklist. This script downloads the blacklist and merges it, then applies this to the alignment.

```{bash, eval=FALSE}

mkdir blacklists
cd blacklists
# Human hg19/GRCh37
wget http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeMapability/wgEncodeDacMapabilityConsensusExcludable.bed.gz --no-check-certificate

# Drosophila dm3
wget http://www.broadinstitute.org/~anshul/projects/fly/blacklist/dm3-blacklist.bed.gz --no-check-certificate  
gunzip dm3-blacklist.bed.gz

### Generate a merged Drosophila/Human blacklist
droso=./dm3-blacklist.bed
homo=./hg19-blacklist.bed

sed "s/^/hs_/" $homo |cut -f1-3 > tmp
sed "s/^/dm_/" $droso > tmp2
cat tmp tmp2 > dmhs-blacklist.bed
rm tmp tmp2

cd ..

bl=../blacklists/dmhs-blacklist.bed

cd ./SLX-8047_dmhs

mkdir ./blacklist_filtered
for f in *.bam
do
	echo $f
	bedtools intersect -v -abam $f -b $bl  > blacklist_filtered/$f
done

### Re-index

cd ./blacklist_filtered
for f in *.bam
do
  samtools index $f
done

```

Peak calling of the reads aligned to the merged genome is then similar to that of a typical experiment.

```{bash, eval=FALSE}

### Run macs on the blacklisted data
mkdir ./SLX-8047_dmhs/peaks
cd ./SLX-8047_dmhs/peaks
control=../blacklist_filtered/SLX-8047.D705_D507.C81G5ANXX.s_1.r_1.fq.gz.bam

for bam in ../blacklist_filtered/*bam
do
  root=`basename $bam .bam`
  macs2 callpeak -t $bam -c $control -f BAM -n $root -g hs
done

```

After aligning the data and calling peaks, we can then split the data back into human and drosophila.

```{bash, eval=FALSE}
# The BAMs and BEDs should be split into Drosophila (for normalization) and into Human (for experimental)

### BAMs, split in two
cd ./SLX-8047_dmhs/blacklist_filtered

mkdir human
mkdir drosophila

for bam in *bam
do
  samtools view -h $bam | grep -v "dm_chr" | sed s/hs_chr/chr/g | samtools view -bS - > human/$bam
  samtools view -h $bam | grep -v "hs_chr" | sed s/dm_chr/chr/g | samtools view -bS - > drosophila/$bam 
done

# Index
cd ./human
echo `pwd`

for bam in *bam
do
  samtools index $bam
done

cd ../drosophila
echo `pwd`

for bam in *bam
do
  samtools index $bam
done

### Peaks BEDs and XLSs, split in two
cd ../../peaks

echo `pwd`

mkdir human
mkdir drosophila

for bed in *peaks.bed
do
  grep -v "dm_chr" $bed | sed s/hs_chr/chr/g > human/$bed
  grep -v "hs_chr" $bed | sed s/dm_chr/chr/g > drosophila/$bed
done

for xls in *peaks.xls
do
  grep -v "dm_chr" $xls | sed s/hs_chr/chr/g > human/$xls
  grep -v "hs_chr" $xls | sed s/dm_chr/chr/g > drosophila/$xls
done

for narrow in *narrowPeak
do
  grep -v "dm_chr" $narrow | sed s/hs_chr/chr/g > human/$narrow
  grep -v "hs_chr" $narrow | sed s/dm_chr/chr/g > drosophila/$narrow
done

```

With the peak files and the separate alignments extracted, we can then write sample sheets. For the normalization we need one for the drosophila aligned reads from each sample and the respective H2av peak file, and one for human reads from each sample and their respsective ER peak files. This forms the bais of the following analysis.

## Load convenience functions

These functions facilitate the normalisation of data.

```{r brundle, results='hide', message=FALSE, warning=FALSE}
source('../package/brundle.R')
```

## Apply settings

```{r}
dbaSummits                <- 200
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
jg.treatedCondition       =  "Fulvestrant"
jg.untreatedCondition     =  "none"

```

## Load control and experimental DiffBind object

As before, to keep file size down, data are provided in Rdata format rather than as raw BAM files.

```{r}
filename<-"Rdata/example_003_SLX-8047_dba_HsDm.rda"
if(!file.exists(filename)){
  dbaExperiment <- jg.getDba(jg.experimentSampleSheet, bRemoveDuplicates=TRUE)
  dbaControl    <- jg.getDba(jg.controlSampleSheet, bRemoveDuplicates=TRUE)
  save(dbaExperiment,dbaControl,file=filename)
} else {
  load(filename)
}

```

```{r}
#Load Sample Ids from control sample sheet.
jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)
```

```{r}

# Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

```

```{r}
#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

```

```{r}
#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

```

```{r}

##Plot showing normalisation calculation (Optional)
  jg.plotNormalization(jg.controlCountsTreated,
                     jg.controlCountsUntreated)
```

```{r}
##Get Normalisation Coefficient
jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)
jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames)

```

```{r}
#Apply coefficent and control factor
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames)
```

```{r}
#Return values to DiffBind and plot normalised result.
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised)

```

```{r}
jg.dba_analysis<-dba.analyze(jg.dba)
dba.plotMA(jg.dba_analysis)

```

## Save results

```{r}
write.csv(dba.report(jg.dba_analysis),file="results/Example_003.csv")
```