### 2017-11-07 Manuscript_Figures.Rmd
Removal of ER proximal CTCF sites

---
title: "Brundle Development"
author: "Andrew Holding"
date: "11/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages=FALSE)
```

## Comparison of simple normalisation methods

```{r}
library(DiffBind)
#setwd("/Volumes/FlyPeaks/FlyPeaks")

hsconsensus<-readRDS("Rdata/015_SLX-12998_hsconsensus.rds")
mmconsensus<-readRDS("Rdata/015_SLX-12998_mmconsensus.rds")
hscounts<-hsconsensus[,-c(1:3)]
mmcounts<-mmconsensus[,-c(1:3)]
### MA RPM in peaks
hsrpm<-apply(hscounts,2,function(x){
  1E6*x/sum(x)
})
mmrpm<-apply(mmcounts,2,function(x){
  1E6*x/sum(x)
})
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM reads in peaks",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}

load("Rdata/012_SLX-12998_aligned.rda")
aligned<-aligned[grep("SLX-12998.D",names(aligned))]
aligned<-sapply(aligned,sum)/1E6
aligned<-aligned[c(4,7,8,9,2,3,6,5)]

hsrpm<-hscounts
for(i in 1:length(hsrpm)){
  hsrpm[i]<-hscounts[i]/aligned[i]
}
mmrpm<-mmcounts
for(i in 1:length(mmrpm)){
  mmrpm[i]<-mmcounts[i]/aligned[i]
}
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}

load("Rdata/012_SLX-12998_nrreads.rda")
nrreads<-nrreads[grep("SLX-12998.D",names(nrreads))]
nrreads<-nrreads/1E6
nrreads<-nrreads[c(4,7,8,9,2,3,6,5)]
hsrpm<-hscounts
for(i in 1:length(hsrpm)){
  hsrpm[i]<-hscounts[i]/nrreads[i]
}
mmrpm<-mmcounts
for(i in 1:length(mmrpm)){
  mmrpm[i]<-mmcounts[i]/nrreads[i]
}
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM total reads",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}
### MA Counts human
# a is treated, b is not treated
M<-apply(hscounts,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hscounts,1,function(x){
  return(log10(sum(x)))
})

par(mar=c(5.1,5.1,4.1,2.1))
plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Raw counts in peaks", ylim=c(-6.25,4.25))
abline(h=0)

```

## MA Plot of H2AV normalization
```{r}
# Read consensus peaks
hsconsensus<-readRDS("Rdata/015_SLX-8047_hsconsensus.rds")
dmconsensus<-readRDS("Rdata/015_SLX-8047_dmconsensus.rds")

# Get counts
hscounts<-hsconsensus[,-c(1:3)]
dmcounts<-dmconsensus[,-c(1:3)]

### MA for Drosophila + Human Counts
# a is treated, b is not treated
Mhs<-apply(hscounts,1,function(x){
  fulvestrant<-mean(x[c(1,3,5,7)])
  untreated<-mean(x[c(2,4,6,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs<-apply(hscounts,1,function(x){
  return(log10(sum(x)))
})
Mdm<-apply(dmcounts,1,function(x){
  fulvestrant<-mean(x[c(1,3,5,7)])
  untreated<-mean(x[c(2,4,6,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Adm<-apply(dmcounts,1,function(x){
  return(log10(sum(x)))
})

plot(Ahs,Mhs,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,2))
points(Adm,Mdm,col="cornflowerblue",pch=20)
abline(h=0)
legend("topright",legend=c("Drosophila","Human"),pch=20,col=c("cornflowerblue","black"))
lm1<-lm(Mdm~Adm)
abline(lm1$coef,col="red")
```

```{r}
### 20th century normalization
# Residuals of the Drosophila fit
lm1<-lm(Mdm~Adm)
intercept<-lm1$coef[1]
angularcoeff<-lm1$coef[2]
MhsFit<-Mhs-(Ahs*angularcoeff)-intercept
MdmFit<-Mdm-(Adm*angularcoeff)-intercept

plot(Ahs,MhsFit,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Counts normalized by H2av Distribution",ylim=c(-6.25,2))
points(Adm,MdmFit,pch=20,col="cornflowerblue")
abline(h=0)
legend("topright",legend=c("Drosophila","Human"),pch=20,col=c("cornflowerblue","black"))
lm1<-lm(MdmFit~Adm)
abline(lm1$coef,col="red")

```

## DeSeq Results

```{r}

source('package/brundle.R')
## Changing font sizes for manuscript

jg.plotDeSeq<-function(ma.df, filename = 'file.name', p = 0.01, title.main = "Differential ChIP",log2fold =0.5, flip=FALSE)
{;

    if (flip == TRUE)
    {
        ma.df$log2FoldChange <- -ma.df$log2FoldChange
    }
    par(mar=c(5.1,5.1,4.1,2.1))
    xyplot(ma.df$log2FoldChange ~ log(ma.df$baseMean, base=10),#xlim=c(0,4),ylim=c(-3,1.25),
           groups=(ma.df$padj < p & abs(ma.df$log2FoldChange) > log2fold & !is.na(ma.df$padj)),
           col=c("black","red"), main=title.main, scales="free", aspect=1, pch=20, cex=0.5,
           ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),
           par.settings=list(axis.text=list(cex=1.5,font=1),par.main.text=list(cex=1.5,font=2),par.xlab.text=list(cex=1.5,font=2), par.ylab.text=list(cex=1.5,font=2)
           )
    );

}

jg.plotDeSeqCombined <- function(jg.controlResultsDeseq,jg.experimentResultsDeseq,title.main,padjX,flip=FALSE)
{
    jg.controlResultsDeseq$group = 'a'
    jg.experimentResultsDeseq$group = 'b'

    if (flip == TRUE)
    {
        jg.controlResultsDeseq$log2FoldChange <- -jg.controlResultsDeseq$log2FoldChange
        jg.experimentResultsDeseq$log2FoldChange <- -jg.experimentResultsDeseq$log2FoldChange
    }

    for (i in 1:length(jg.experimentResultsDeseq$group)) {
        if (!is.na(jg.experimentResultsDeseq$padj[i]) & !is.na(jg.experimentResultsDeseq$log2FoldChange[i]) & jg.experimentResultsDeseq$padj[i] < padjX & jg.experimentResultsDeseq$log2FoldChange[i] < 0) {
            jg.experimentResultsDeseq$group[i] <- 'd'
        }
        else if (!is.na(jg.experimentResultsDeseq$padj[i]) & !is.na(jg.experimentResultsDeseq$log2FoldChange[i]) & jg.experimentResultsDeseq$padj[i] < padjX & jg.experimentResultsDeseq$log2FoldChange[i] > 0) {
            jg.experimentResultsDeseq$group[i] <- 'c'
        }
    }

    for (i in 1:length(jg.controlResultsDeseq$group)) {
        if (!is.na(jg.controlResultsDeseq$padj[i]) & !is.na(jg.controlResultsDeseq$log2FoldChange[i]) & jg.controlResultsDeseq$padj[i] < padjX & jg.controlResultsDeseq$log2FoldChange[i] < 0) {
            jg.controlResultsDeseq$group[i] <- 'f'
        }
        else if (!is.na(jg.controlResultsDeseq$padj[i]) & !is.na(jg.controlResultsDeseq$log2FoldChange[i]) & jg.controlResultsDeseq$padj[i] < padjX & jg.controlResultsDeseq$log2FoldChange[i] > 0) {
            jg.controlResultsDeseq$group[i] <- 'e'
        }
    }

    full.res = rbind(jg.controlResultsDeseq, jg.experimentResultsDeseq)
    par(mar=c(5.1,5.1,4.1,2.1))
    xyplot(full.res$log2FoldChange ~ log(full.res$baseMean, base=10), data = full.res,
           groups=full.res$group,
           col=c("grey40","grey80",  "#5480ff", "#ff5454", "#08298a","#750505"),
           ylab = expression('log'[2]*' Differential ChIP'),
           xlab = expression("log"[10]~"Mean of Normalized Counts"),
           aspect=1.0,
           pch=16,
           cex=0.5,
           main=title.main,
           scales=list(x=list(cex=1.5, relation = "free"), y =list(cex=1.5, relation="free")),
           between=list(y=0.5, x=0.5),
           auto.key = TRUE,
           par.settings=list(axis.text=list(cex=1.5,font=1),par.main.text=list(cex=1.5,font=2),par.xlab.text=list(cex=1.5,font=2), par.ylab.text=list(cex=1.5,font=2)),
           key=list(corner=c(1,0),
                    cex=1.0,
                    points=list(col=c( "gray80","gray40", "#ff5454", "#5480ff", "#750505", "#08298a","white"), pch=20),
                    text=list(c("Target Binding","Control Binding", "Target Decreased","Target Increased","Control Decreased","Control Increased", " "))
           ))

}

######################
#
# Settings
#
######################

setwd("/Volumes/FlyPeaks/FlyPeaks")
dbaSummits                <- 200
jg.controlMinOverlap      <- 5
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14229_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14229_hs_ER_DBA.csv"

filename<-"Rdata/028_SLX-14229_dba_human_ER_CTCF.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)

  jg.plotDeSeq(jg.controlResultsDeseq,
               p=0.01, 
               title.main="Fold-change in CTCF binding",
               flip=T
               )

```

```{r}
jg.plotDeSeq(jg.experimentResultsDeseq,
             p=0.01,
             title.main="Fold-change in ER binding",
             flip=T
             )
```

#CTCF Binding

```{r}

jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)

 jg.plotDeSeq(jg.experimentResultsDeseqInternal,
               title.main="Fold-change in ER binding (no correction)",
               p=0.01,
               flip=T
               )
```

#H2av With Diff Bind
```{r}
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_hs.csv"
filename<-"Rdata/029_SLX-8047_dba_human_drosophila.rda"

load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)

jg.experimentDeSeqInternal<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions, NULL)
jg.experimentResultsDeseqInternal   = results(jg.experimentDeSeqInternal)

dbaExperimentCorrected<-jg.correctDBASizeFactors(dbaExperiment,
                                                 jg.controlSizeFactors
                                                 )

dbaExperimentAnalysis<-dba.analyze(dbaExperiment)
  dba.plotMA(dbaExperimentAnalysis)

```

```{r}
  dbaExperimentAnalysisCorrected<-dba.analyze(dbaExperimentCorrected)
  dba.plotMA(dbaExperimentAnalysisCorrected)
```
```{r}

filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
  load(filename)
  jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"
jg.treatedCondition       =  "Fulvestrant"
jg.untreatedCondition     =  "none"

#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

jg.plotNormalization(jg.controlCountsTreated, jg.controlCountsUntreated)
```
```{r}

jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX8047_dm.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX8047_consensus.csv"

filename<-"Rdata/029_SLX-8047_SLX14229_dba_human_drosophila.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq_SLX8047   = results(jg.experimentDeSeq)

#Now run for SLX-14229

jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14229_hs_CTCF_DBA.csv"
#Using the same consensus peak set from above 
jg.experimentSampleSheet  <- "samplesheet_SLX14229_hs_ER_consensus.csv"

filename<-"Rdata/029_SLX-8047_SLX-14229_dba_human_ER_CTCF_consensus.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq_SLX14229   = results(jg.experimentDeSeq)

library(latticeExtra)

ma.df<-jg.experimentResultsDeseq_SLX14229

a<-    xyplot((-ma.df$log2FoldChange+0.2) ~ log(ma.df$baseMean, base=10),
              panel=function(...) {
                  panel.xyplot(...)
                  panel.abline(h=0, lty = "dotted", col = "black")
                  panel.segments(3.0,0,1,1.6, col="red",lwd=2)
                  panel.segments(3.0,0,1,-1.6, col="red",lwd=2)
                  panel.segments(1,1.6,1,-1.6, col="red",lwd=2)
              },
           col=c("deepskyblue4"), ylim=c(-6,2), main="Comparision between CTCF and H2av", scales="free", aspect=1, pch=20, cex=0.5,
           ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),
           par.settings=list(par.main.text=list(cex=1.57,font=2),axis.text=list(cex=1.57,font=1),par.xlab.text=list(cex=1.57,font=2), par.ylab.text=list(cex=1.57,font=2)));

ma.df_2<-jg.experimentResultsDeseq_SLX8047
#Note this has been manually addjusted to be more illustrative.
b<-    xyplot(-(ma.df_2$log2FoldChange*1.1-0.1) ~ (log(ma.df_2$baseMean, base=10)*1.1+0.2),
              panel=function(...) {
                  panel.xyplot(...)
                  panel.abline(h=0, lty = "dotted", col = "black")
                  panel.segments(3.0,0,1,1.6, col="red",lwd=2)
                  panel.segments(3.0,0,1,-1.6, col="red",lwd=2)
                  panel.segments(1,1.6,1,-1.6, col="red",lwd=2)
              },
              col=c("palegreen3"), main="Comparision between CTCF and H2av", scales="free", aspect=1, pch=20, cex=0.5,
              ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),
              par.settings=list(par.xlab.text=list(cex=1.1,font=2), par.ylab.text=list(cex=1.1,font=2)));

a + as.layer(b)

```
```{r}

plot(jg.experimentResultsDeseq_SLX14229$log2FoldChange,
     jg.experimentResultsDeseq_SLX8047$log2FoldChange,
     pch=20,
     cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,
     main="Comparision of fold change between datasets",
     xlab=expression("log"[2]~" ER ChIP fold change CTCF normalised"),
     ylab=expression("log"[2]~" ER ChIP fold change H2av normalised"))

```

```{r}

jg.plotScatter<-function(peaks,samples,yaxis,samplesCombinations,noYLabel=FALSE)
{
    peaks$count <- 1
    peaks_aggregate <- aggregate(count ~ ., peaks, FUN = sum)
    colormap<-rainbow(255,start=0,end=4/6)
    colormap<-rev(colormap)
    peaks_aggregate$Col <- colormap[as.numeric(cut(log(peaks_aggregate$count,2),breaks=255))]
    peaks_aggregate<-peaks_aggregate[order(peaks_aggregate$count),]
    plot(log(peaks_aggregate[,c(1,2)]), col=peaks_aggregate$Col,pch=20, cex=0.3, xlim=c(0,7),ylim=c(0,7),axes=FALSE)
    if (samples > 1|| noYLabel==TRUE) {Axis(side=2, labels=FALSE)} else {
        Axis(side=2, labels=TRUE)
        mtext("log(Counts)", side=2, line=3)
    }
    if (yaxis !=TRUE ) {Axis(side=1, labels=FALSE)} else {
        Axis(side=1, labels=TRUE,ylab="Log(Counts)")
        mtext("log(Counts)", side=1, line=3)
    }
    text(1.5,6.5,paste0(samplesCombinations[,samples][1],"-",samplesCombinations[,samples][2]),cex=1.5)
    sampleCor<-cor(log(peaks[1]),log(peaks[2]))
    text(2,6.0,paste0("Correlation = ",signif(sampleCor,3)),cex=1.0)
        abline(a=0,b=1,col="grey")
    return(cor(log(peaks_aggregate[,c(1,2)])))
}

jg.plotMargins<-function(a,b)
{
    par(mar=c(0.5, 0.5, 0.2, 0.2), mfrow=c(a,b),
        oma = c(4, 4, 0.2, 0.2))
}
jg.plotPanelsPlot<-function(treatmentNames,controlNames,dbaExperiment,noYLabel=FALSE)
{

    jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
    relativeLibrarySize<-jg.getLibrarySize(dbaExperiment,
                                            jg.experimentPeakset)

    samplesCombinations<-combn(controlNames,2)
    for (samples in seq(1,length(samplesCombinations)/2))  {
        peaksLibrarySize<-relativeLibrarySize[samplesCombinations[,samples]]
        peaks<-jg.experimentPeakset[samplesCombinations[,samples]]
        peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
        peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
        jg.plotScatter(peaks,samples, F,samplesCombinations,noYLabel)
    }
    samplesCombinations<-combn(treatmentNames,2)
    for (samples in seq(1,length(samplesCombinations)/2))  {
        peaksLibrarySize<-relativeLibrarySize[samplesCombinations[,samples]]
        peaks<-jg.experimentPeakset[samplesCombinations[,samples]]
        peaks[1]<-round(peaks[1]/peaksLibrarySize[1])
        peaks[2]<-round(peaks[2]/peaksLibrarySize[2])
        jg.plotScatter(peaks,samples,T,samplesCombinations,noYLabel)
    }
 }

jg.getLibrarySize<-function(dbaExperiment,jg.experimentPeakset) {
    librarySize<-as.numeric(dbaExperiment$class["Reads",])
    relativeLibrarySize<-librarySize/max(librarySize)
    names(relativeLibrarySize)<-colnames(jg.experimentPeakset[c(-1:-3)])
    return(relativeLibrarySize)
}

jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"

filename<-"Rdata/048_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

controlNames<-c("1a","2a","3a")
treatmentNames<-c("1b","2b","3b")

filename<-"Rdata/026_SLX-8047_dba.counts_human.rda"
load(filename)
dbaExperiment_SLX8047<-dba

controlNames_SLX8047<-c("1b","2b","3b","4b")
treatmentNames_SLX8047<-c("1a","2a","3a","4a")

filename<-"Rdata/047_SLX-12998_dba_human.rda"
load(filename)
dbaExperiment_SLX12998<-dbaFullExperiment

controlNames_SLX12998<-c("1b","2b","3b","4b")
treatmentNames_SLX12998<-c("1a","2a","3a","4a")

jg.plotMargins(2,3)
par(mfcol=c(2,3))

#SLX-14438
#2a-3a and 1b-3b
controlNames<-c("2a","3a")
treatmentNames<-c("1b","3b")
jg.plotPanelsPlot(treatmentNames,
                  controlNames,
                  dbaExperiment)

#SLX-8047
#3b-4b and 1a-4a
controlNames<-c("3b","4b")
treatmentNames<-c("1a","4a")
jg.plotPanelsPlot(treatmentNames,
                  controlNames,
                  dbaExperiment_SLX8047, TRUE)

#SLX-12998
#3b-4b and 1a-4a
controlNames<-c("3b","4b")
treatmentNames<-c("1a","4a")
jg.plotPanelsPlot(treatmentNames,
                  controlNames,
                  dbaExperiment_SLX12998,TRUE)
```

```{r}
filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

dba.plotVenn(dbaControl,dbaControl$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", main="CTCF Peaks Fulvestrant")
dba.plotVenn(dbaControl,dbaControl$masks$none, label1="Rep1", label2="Rep2", label3="Rep3", main="CTCF Peaks Control")
dba.plotVenn(dbaExperiment,dbaExperiment$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", main="ER Peaks Fulvestrant")
dba.plotVenn(dbaExperiment,dbaExperiment$masks$none,label1="Rep1", label2="Rep2", label3="Rep3", main="ER Peaks Control")
```
```{r}

filename<-"Rdata/026_SLX-8047_dba_drosophila.rda"
load(filename)
filename<-"Rdata/026_SLX-8047_dba_human.rda"
load(filename)

dba.plotVenn(dba_dm,dba_dm$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="H2av Peaks Fulvestrant")
dba.plotVenn(dba_dm,dba_dm$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="H2av Peaks Control")
dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="ER Peaks Fulvestrant")
dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="ER Peaks Control")
```

```{r}
filename<-"Rdata/014_SLX-12998_dba_mouse.rda"
load(filename)

dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="Mouse ER Peaks Fulvestrant")
dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="Mouse ER Peaks Control")

filename<-"Rdata/015_SLX-12998_dba_human.rda"
load(filename)

dba.plotVenn(dba,dba$masks$Fulvestrant, label1="Rep1", label2="Rep2", label3="Rep3", label4="Rep4",main="ER Peaks Fulvestrant")
dba.plotVenn(dba,dba$masks$none,label1="Rep1", label2="Rep2", label3="Rep3",label4="Rep4", main="ER Peaks Control")
```

```{r}

# Read consensus peaks
hsconsensus<-readRDS("Rdata/015_SLX-12998_hsconsensus.rds")
mmconsensus<-readRDS("Rdata/015_SLX-12998_mmconsensus.rds")

# Get counts
hscounts<-hsconsensus[,-c(1:3)]
mmcounts<-mmconsensus[,-c(1:3)]
head(hscounts)
head(mmcounts)

### MA RPM in peaks
hsrpm<-apply(hscounts,2,function(x){
  1E6*x/sum(x)
})
mmrpm<-apply(mmcounts,2,function(x){
  1E6*x/sum(x)
})
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})

load("Rdata/012_SLX-12998_aligned.rda")
aligned<-aligned[grep("SLX-12998.D",names(aligned))]
aligned<-sapply(aligned,sum)/1E6
aligned<-aligned[c(4,7,8,9,2,3,6,5)]

hsrpm<-hscounts
for(i in 1:length(hsrpm)){
  hsrpm[i]<-hscounts[i]/aligned[i]
}
mmrpm<-mmcounts
for(i in 1:length(mmrpm)){
  mmrpm[i]<-mmcounts[i]/aligned[i]
}
M<-apply(hsrpm,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A<-apply(hsrpm,1,function(x){
  return(log10(sum(x)))
})
plot(A,M,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4.25))
abline(h=0)

```

```{r}

### MA for Mouse + Human Counts
# a is treated, b is not treated
Mhs<-apply(hscounts,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs<-apply(hscounts,1,function(x){
  return(log10(sum(x)))
})
Mmm<-apply(mmcounts,1,function(x){
  fulvestrant<-mean(x[c(1,2,4,5)])
  untreated<-mean(x[c(3,6,7,8)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Amm<-apply(mmcounts,1,function(x){
  return(log10(sum(x)))
})

plot(Ahs,Mhs,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="RPM aligned reads",ylim=c(-6.25,4))
points(Amm,Mmm,pch=20,col="darkolivegreen3")
abline(h=0)
legend("topright",legend=c("Mouse","Human"),pch=20,col=c("darkolivegreen3","black"))
lm1<-lm(Mmm~Amm)
abline(lm1$coef,col="red")

```

```{r}

# Residuals of the Drosophila fit
lm1<-lm(Mmm~Amm)
intercept<-lm1$coef[1]
angularcoeff<-lm1$coef[2]
MhsFit<-Mhs-(Ahs*angularcoeff)-intercept
MmmFit<-Mmm-(Amm*angularcoeff)-intercept

plot(Ahs,MhsFit,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"),main="Counts normalized by Mouse Distribution",ylim=c(-16,2))
points(Amm,MmmFit,pch=20,col="darkolivegreen3")
abline(h=0)
legend("topright",legend=c("Mouse","Human"),pch=20,col=c("darkolivegreen3","black"))
lm1<-lm(MmmFit~Amm)
abline(lm1$coef,col="red")
```

```{r}
hsconsensus_ER<-readRDS("Rdata/018_SLX-14229_hsconsensus_ER.rds")
hsconsensus_CTCF<-readRDS("Rdata/018_SLX-14229_hsconsensus_CTCF.rds")
hscounts_ER<-hsconsensus_ER[,-c(1:3)]
hscounts_CTCF<-hsconsensus_CTCF[,-c(1:3)]

load("Rdata/012_SLX-14229_aligned.rda")
aligned<-aligned[grep("SLX-14229.D",names(aligned))]
aligned<-sapply(aligned,sum)/1E6

#remove inputs/controls to match samples here
aligned<-aligned[c(1,2,5,7,8,10)]

hsrpm_ER<-hscounts_ER
for(i in 1:length(hsrpm_ER)){
  hsrpm_ER[i]<-hscounts_ER[i]/aligned[i]
}
M_RPM_ER<-apply(hsrpm_ER,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A_RPM_ER<-apply(hsrpm_ER,1,function(x){
  return(log10(sum(x)))
})
plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
abline(h=0)

```

```{r}

hsrpm_CTCF<-hscounts_CTCF
for(i in 1:length(hsrpm_CTCF)){
  hsrpm_CTCF[i]<-hscounts_CTCF[i]/aligned[i]
}
M_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
A_RPM_CTCF<-apply(hsrpm_CTCF,1,function(x){
  return(log10(sum(x)))
})

plot(A_RPM_ER,M_RPM_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="RPM aligned reads")
points(A_RPM_CTCF,M_RPM_CTCF,pch=20,col="gray")
lm1<-lm(M_RPM_CTCF~A_RPM_CTCF)
abline(lm1$coef,col="blue")
abline(h=0)

```

```{r}

Mhs_CTCF<-apply(hscounts_CTCF,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs_CTCF<-apply(hscounts_CTCF,1,function(x){
  return(log10(sum(x)))
})

Mhs_ER<-apply(hscounts_ER,1,function(x){
  fulvestrant<-mean(x[c(2,4,5)])
  untreated<-mean(x[c(1,3,6)])
  fc<-mean(fulvestrant)/mean(untreated)
  log2fc<-log2(fc)
  return(log2fc)
})
Ahs_ER<-apply(hscounts_ER,1,function(x){
  return(log10(sum(x)))
})

lm1<-lm(Mhs_CTCF~Ahs_CTCF)
intercept<-lm1$coef[1]
angularcoeff<-lm1$coef[2]
MhsFit_ER<-Mhs_ER-(Ahs_ER*angularcoeff)-intercept
MhsFit_CTCF<-Mhs_CTCF-(Ahs_CTCF*angularcoeff)-intercept
plot(Ahs_ER,MhsFit_ER,cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5,pch=20,ylab=expression("log"[2]~"ChIP fold change"), xlab=expression("log"[10]~"Mean of Normalized Counts"), main="Normalised to Human CTCF")
points(Ahs_CTCF,MhsFit_CTCF,pch=20,col="gray")

abline(h=0)
legend("topright",legend=c("Human ER",  "Human CTCF"),pch=20,col=c("black","gray"), cex=1)

```
```{r}
filename<-"Rdata/012_SLX-12998_nrreads.rda"
load(filename)

filename<-"Rdata/012_SLX-12998_aligned.rda"
load(filename)

nrreads<-nrreads[1:10]#Remove lost reads
aligned<-aligned[1:10]

### Prepare relative alignments
samples<-names(nrreads)
toplot<-matrix(nrow=length(samples),ncol=5)
colnames(toplot)<-c("MmTot","HsTot","Tot","MmRel","HsRel")
rownames(toplot)<-samples

mms<-sapply(aligned,function(x){
  sum(x[grep("mm_",names(x))])
})
hss<-sapply(aligned,function(x){
  sum(x[grep("hs_",names(x))])
})
toplot[,1]<-mms
toplot[,2]<-hss
toplot[,3]<-mms+hss
toplot[,4]<-mms/(mms+hss)
toplot[,5]<-hss/(mms+hss)

toplot<-toplot[grep('SLX-12998.D', sort(rownames(toplot)), value=TRUE),]

## Barplot of relative alignments
# Convert sample names to informative ones

rownames(toplot)<-c(
  "Input ICI",
  "2 ICI",
  "1 Control",
  "4 ICI",
  "2 Control",
  "3 Control",
  "1 ICI",
  "4 Control",
  "3 ICI",
  "Input Control"
)

# Sort by total reads aligned

toplot<-toplot[order(-toplot[,3]),]

par(las=2,mar=c(7,4,3,5))
bp<-barplot(t(toplot[,4:5]),beside=TRUE,main="Relative Read Alignment in samples",
            ylab="Fraction of total aligned",col=c("royalblue2","tomato"),ylim=c(-0.05,1.05))
grid()
legend("topright",legend=c("Mouse","Human"),col=c("royalblue2","tomato"),pch=15)
bpx<-apply(bp,2,mean)

raxis<-(toplot[,3])/(max(toplot[,3]))
lines(bpx,raxis,type="b",lwd=3,pch=16)
axis(4,at=seq(0,1,0.2),
     labels=round((seq(0,1,0.2) * max(toplot[,3])/10**6))
     )
par(las=3)
mtext("Reads aligned (Millions)",side=4,line=4)

```
```{r}
filename<-"Rdata/015_SLX-14229_dbacounts_human.rda"
load(filename)
plot(dbacounts)

filename<-"Rdata/018_SLX-14229_dba.counts_human_CTCF.rda"
load(filename)
plot(dba)

filename<-"Rdata/018_SLX-14229_dba.counts_human_ER.rda"
load(filename)
plot(dba)
```

```{r}
jg.controlSampleSheet     <- "samplesheet/samplesheet_SLX14438_hs_CTCF_DBA.csv"
jg.experimentSampleSheet  <- "samplesheet/samplesheet_SLX14438_hs_ER_DBA.csv"

filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)
jg.correctionFactor<-jg.getCorrectionFactor(jg.experimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames)

#Apply coefficent and control factor
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames)
#Return values to Diffbind and plot normalised result.
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised)
jg.dba_analysis<-dba.analyze(jg.dba)

dba.plotMA(jg.dba_analysis,bFlip=TRUE, cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
dba_analysis<-dba.analyze(dbaExperiment)
dba.plotMA(dba_analysis,bFlip=TRUE,cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)
#Note exact numbers change depending on the machine DiffBind is run.

```
```{r}
filename<-"Rdata/040_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Convert Peakset to DeSeq Workflow
jg.controlPeaksetDeSeq<-jg.convertPeakset(jg.controlPeakset)

#Establish size factors directly from Control data
jg.controlSizeFactors = estimateSizeFactorsForMatrix(jg.controlPeaksetDeSeq)

#Get conditions dataframe for DeSeq
jg.conditions <- read.csv(file=jg.controlSampleSheet, header=TRUE, sep=",")['Condition']

#Run DeSeq on control
jg.controlDeSeq<-jg.runDeSeq(jg.controlPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.controlResultsDeseq   = results(jg.controlDeSeq)

#Repeat for experimental conditions

#Convert experiment Peakset to DeSeq Workflow
jg.experimentPeaksetDeSeq<-jg.convertPeakset(jg.experimentPeakset)

#Run DeSeq on experiment
jg.experimentDeSeq<-jg.runDeSeq(jg.experimentPeaksetDeSeq, jg.conditions,jg.controlSizeFactors)
jg.experimentResultsDeseq   = results(jg.experimentDeSeq)
dbaExperimentCorrected<-jg.correctDBASizeFactors(dbaExperiment,
                                                 jg.controlSizeFactors
                                                 )

dbaExperimentAnalysisCorrected<-dba.analyze(dbaExperimentCorrected)
  dba.plotMA(dbaExperimentAnalysisCorrected,bFlip=TRUE,cex.lab=1.5, cex.axis=1.5, cex.main=1.25, cex.sub=1.5)

  #Again numbers similar but not exact due to DiffBind
```
```{r}
filename<-"Rdata/032_SLX-14229_dba_human_ER_CTCF_normalised.rda"
load(filename)

filename<-"Rdata/032_SLX-8047_dba_normalised.rda"
load(filename)

jg.dba_analysis_SLX8047<-dba.analyze(jg.dba_SLX8047)
jg.dba_analysis_SLX14229<-dba.analyze(jg.dba_SLX14229)

nearestPeak <- nearest(dba.report(jg.dba_analysis_SLX8047, th=1),dba.report(jg.dba_analysis_SLX14229, th=1))
SLX8047_fc<-dba.report(jg.dba_analysis_SLX8047,bFlip=TRUE, th=1)$Fold
SLX14229_fc<-dba.report(jg.dba_analysis_SLX14229,th=1)[nearestPeak]$Fold

smoothScatter(SLX8047_fc,SLX14229_fc,pch=20, col="#00000055",
              xlab="Fold change normalized using Drosophila/H2av spike-in",
              ylab="Fold change normalized using CTCF control peaks",
              main="Comparision of normalisation methods",
              cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
abline(h=0, col="grey",lty=2)
abline(v=0, col="grey",lty=2)
coef<-lm( SLX8047_fc ~ SLX14229_fc + 0 )
abline(coef, col="navyblue")
text(3.25,-0.25, expression( "gradient = 0.96, r = 0.557, p-value < 2.2 x" ~ 10^{-16}), cex=1.25)

```
```{r}

```

```{r}
filename<-"Rdata/047_SLX-14438_dba_human_ER_CTCF.rda"
load(filename)

#Load Sample Ids from control sample sheet.
jg.sampleIds <- jg.getSampleIds(jg.controlSampleSheet)

## Extract Peak set from DiffBind
jg.experimentPeakset <- jg.dbaGetPeakset(dbaExperiment)
jg.controlPeakset    <- jg.dbaGetPeakset(dbaControl)

#Get counts for each condition
jg.controlCountsTreated<-jg.getControlCounts(jg.controlPeakset, 
                                             jg.controlSampleSheet,
                                             jg.treatedCondition)
jg.controlCountsUntreated<-jg.getControlCounts(jg.controlPeakset,
                                               jg.controlSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.controlCountsUntreated)
jg.treatedNames   <- names(jg.controlCountsTreated)

 ##Get Normalization Coefficient
jg.coefficient<-jg.getNormalizationCoefficient(jg.controlCountsTreated,
                                               jg.controlCountsUntreated)

#As we're going to normalise a second data set we don't need to use a DiffBind correction factor
jg.correctionFactor<-1

#Apply coefficent and control factor
jg.experimentPeaksetNormalised<-jg.applyNormalisation(jg.experimentPeakset,
                                                      jg.coefficient,
                                                      jg.correctionFactor,
                                                      jg.treatedNames
)

#Return values to Diffbind object
jg.dba <- DiffBind:::pv.resetCounts(dbaExperiment,
                                    jg.experimentPeaksetNormalised
)

#jg.dba is now normalised a list of ER peaks and their expected ratio

#####
#
#Rerun the normalisation code to get the ER ratio
#
######

jg.normalisedExperimentPeakset <- jg.dbaGetPeakset(jg.dba)

#Get counts for each condition
jg.ERCountsTreated<-jg.getControlCounts(jg.normalisedExperimentPeakset, 
                                             jg.experimentSampleSheet,
                                             jg.treatedCondition)
jg.ERCountsUntreated<-jg.getControlCounts(jg.normalisedExperimentPeakset,
                                          jg.experimentSampleSheet,
                                               jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.ERCountsTreated)
jg.treatedNames   <- names(jg.ERCountsUntreated)

##Establish normalised ER ratio as we know it isn't 1

jg.ERratio<-jg.getNormalizationCoefficient(jg.ERCountsTreated,
                                               jg.ERCountsUntreated)

########
#
# Apply to ER only data
#
########

jg.EROnlyExperimentSampleSheet <- "samplesheet/samplesheet_SLX8047_hs.csv"

filename<-"Rdata/047_SLX-SLX-8047_dba_human_ER.rda"
load(filename)

#Get peakset

jg.EROnlyExperimentPeakset <- jg.dbaGetPeakset(dbaERonlyExperiment)

#Get counts for each condition
jg.EROnlyCountsTreated<-jg.getControlCounts(jg.EROnlyExperimentPeakset, 
                                            jg.EROnlyExperimentSampleSheet,
                                        jg.treatedCondition)
jg.EROnlyCountsUntreated<-jg.getControlCounts(jg.EROnlyExperimentPeakset,
                                              jg.EROnlyExperimentSampleSheet,
                                          jg.untreatedCondition)

#Get sample names for conditions
jg.untreatedNames <- names(jg.EROnlyCountsTreated)
jg.treatedNames   <- names(jg.EROnlyCountsUntreated)

jg.EROnlyExperimentRatio<-jg.getNormalizationCoefficient(jg.EROnlyCountsTreated,
                                           jg.EROnlyCountsUntreated)

#Create new ratio 
jg.ERNormalisationRatio<-jg.ERratio/jg.EROnlyExperimentRatio

g.correctionFactor<-jg.getCorrectionFactor(jg.EROnlyExperimentSampleSheet,
                                            jg.treatedNames,
                                            jg.untreatedNames
)

jg.ERexperimentPeaksetNormalised<-jg.applyNormalisation(jg.EROnlyExperimentPeakset,
                                                      jg.ERNormalisationRatio,
                                                      jg.correctionFactor,
                                                      jg.treatedNames
)

#Return values to Diffbind object
dbaERonlyExperimentNormalised <- DiffBind:::pv.resetCounts(dbaERonlyExperiment,
                                    jg.ERexperimentPeaksetNormalised
)

jg.dba_analysisNormalised<-dba.analyze(dbaERonlyExperimentNormalised)
false_postive_normalised<-dba.report(jg.dba_analysisNormalised, th=0.05)
false_postive_normalised<-length(false_postive_normalised[false_postive_normalised$Fold>0])

jg.dba_analysis<-dba.analyze(dbaERonlyExperiment)

load(file="Rdata/027_SLX-8047_dba_report.Rda")
dba.SLX8047<-dba.analyze(dba.SLX8047)

peaks<-as.character(sort(as.numeric(row.names(as.data.frame(dba.report(dba.SLX8047, th=1))))))

plot(
    dba.report(dba.SLX8047, th=1)[peaks]$Fold,
    dba.report(jg.dba_analysisNormalised,th=1)[peaks]$Fold,
    pch=20,
    xlab="Log(FoldChange) in ER Binding Normalised to H2av",
    ylab="Log(FoldChange) in ER Binding Cross-Normalised to CTCF",
    cex=0.05,
    main="Comparision of Fold-Change between\nCross-Normalised to Xenogenic Spike-in"
)
```